Quality Definitions of Qualities

### 2019-11-20 StateOfPractice/QDefOfQualities/QDefOfQualities.tex
Add resuability and understandability from Ghezzi1991 text boot
Review and comments on several qualities, in prep for group meeting on Nov 20, 2019

\begin{defn}
  ``The capability of a software product to be installed in a specified
  environment'' \citep{berander2005software}.
\end{defn}

@@

\begin{defn}
  The degree to which a system or component facilitates the establishment of
  test criteria and the performance of tests to determine whether those criteria
  have been met \citep{IEEEStdGlossarySET1990}.
\end{defn}

@@

\begin{defn}
  Validatability of a software is the degree of ease in validating (checking)
  that software meets user needs.
\end{defn}

@@

\subsection{\notdone{Robustness} \pmi{owner}}
\begin{defn}
  The degree to which a system or component can function correctly in the
  presence of invalid inputs or stressful environmental conditions
  \citep{IEEEStdGlossarySET1990}.
\end{defn}
\begin{defn}
  The quality can be further informally refined as the ability of a software to
  keep an acceptable behaviour, expressed in terms of robustness requirements, in
  spite of exceptional or unforeseen execution conditions (such as the
  unavailability of system resources, communication failures, invalid or
  stressful inputs, etc.) \citep{fernandez2005model}.
\end{defn}
\begin{defn}
  Code possesses the characteristic of robustness to the extent that it can
  continue to perform despite some violation of the assumptions in its
  specification \citep{boehm2007software}.
\end{defn}
\begin{defn} \label{RobustnessDefnSelected}
  A program is robust if it behaves ``reasonably'', even in circumstances that
  were not anticipated in the requirements specification - for example, when it
  encounters incorrect input data or some hardware malfunction
  \citep{ghezzi1991fundamentals}.
\end{defn}

@@

Definition \ref{RobustnessDefnSelected} rephrased: Software possesses the
characteristic of robustness if it behaves ``reasonably'' in circumstances that
were not anticipated in the requirements specification or are in violation of
it.

@@

Combined definition \ref{PerformanceDefnSelected} and
\ref{PerformanceDefnSelected2}: The degree to which a system or component
accomplishes its designated functions within given constraints, such as speed
(database response times, for instance), throughput ( transactions per second),
capacity (con-current usage loads), and timing (hard real-time demands).

@@

This definition includes all important categories of performance.

@@

\begin{defn} \label{MaintainabilityDefnSelected1}
The ease with which a software system or component can be modified to correct
faults, improve performance or other attributes, or adapt to a changed
environment \citep{IEEEStdGlossarySET1990}. 
\end{defn}\
\begin{defn}
ISO/IEC 25010 refers to maintainability as the degree of effectiveness and
efficiency with which a product or system can be modified by the intended
maintainers \citep{ISO/IEC25010}.
\end{defn}
\begin{defn}
Effort required to locate and fix an error in a program
\citep{pressman2005software}.
\end{defn}
\begin{defn}
A set of attributes that bear on the effort needed to make specified
modifications (which may include corrections, improvements, or adaptations of
software to environmental changes and changes in the requirements and functional
specifications)\citep{pfleeger2006software}.
\end{defn}
\begin{defn}
We will view maintainability as two separate qualities: repairability and
evolvability. Software is repairable if it allows the fixing of defects; it is
evolvable if it allows changes that enable it to satisfy new requirements
\citep{ghezzi1991fundamentals}.
\end{defn}
\begin{defn} \label{MaintainabilityDefnSelected2}
Code possesses the characteristic of maintainability to the extent that it
facilitates updating to satisfy new requirements or to correct deficiencies
\citep{boehm2007software}.
\end{defn}

\begin{defn}
  ``The effort required to locate and fix an error in an operational program''
  \citep{McCallEtAl1977}. (As summarized in \citet{VanVliet2000}.) \wss{It looks
    like \citet{pressman2005software} was using \citet{McCallEtAl1977} for their
    definition.}
\end{defn}

@@

This definition includes all potential reasons to modify the software.

@@

\begin{defn}
  ``The extent to which a program (or parts thereof) can be reused in other
  applications'' \citep{McCallEtAl1977}. (As summarized in
  \citet{VanVliet2000}.)
\end{defn}

@@

\subsection{\notdone{Portability} \pmi{owner}}
\begin{defn}
The ease with which a system or component can be transferred from one hardware
or software environment to another \citep{IEEEStdGlossarySET1990}.
\end{defn}
\begin{defn}
An application is portable across a class of environments to the degree that the
effort required to transport and adapt it to a new environment in the class is
less than the effort of redevelopment \citep{mooney1990strategies}.
\end{defn}
\begin{defn} \label{PortabilityDefnSelected}
Effort required to transfer the program from one hardware and/or software system
environment to another \citep{pressman2005software}.
\end{defn}
\begin{defn}
A set of attributes that bear on the ability of software to be transferred from
one environment to another (including the organizational, hardware, of software
environment)\citep{pfleeger2006software}.
\end{defn}
\begin{defn}
Code possesses the characteristic of portability to the extent that it can be
operated easily and well on computer configurations other than its current
one. This implies that special function features, not easily available at other
facilities, are not used, that standard library functions and subroutines are
selected for universal applicability, and so on \citep{boehm2007software}.
\end{defn}
\begin{defn}
Portability refers to the ability to run a system on different hardware
platforms \citep{ghezzi1991fundamentals}.
\end{defn}
\begin{defn}
  ``The effort required to transfer a program from one hardware and/or software
  environment to another'' \citep{McCallEtAl1977}. (As summarized in
  \citet{VanVliet2000}.)
\end{defn}

@@

Definition \ref{PortabilityDefnSelected} rephrased: Effort required to transfer
a program between system environments.

@@

\begin{defn}
Interoperability is the ability of two or more systems or components to exchange
information and to use the information that has been exchanged
\citep{IEEEComputerDictionary1991}.
\end{defn}
\begin{defn}
\label{InteroperabilitySelected}
The degree to which two or more systems, products or components can exchange
information and use the information that has been exchanged
\citep{ISO/IEC25010}.
\end{defn}
\begin{defn}
The ability of a system to coexist and cooperate with other systems.
\citep{ghezzi1991fundamentals}.
\end{defn}
\begin{defn}
The capability to communicate, execute programs, and transfer data among various
functional units in a manner that requires the user to have little or no
knowledge of the unique characteristics of those units
\citep{ISO/IEC/IEEE24765}.
\end{defn}
\begin{defn}Interoperability is a characteristic of a product or system, whose
interfaces are completely understood, to work with other products or systems,
present or future, in either implementation or access, without any restrictions
\citep{AFUL2019}.
\end{defn}
\begin{defn}
Interoperability is the ability of different information systems, devices and
applications (‘systems’) to access, exchange, integrate and cooperatively use
data in a coordinated manner, within and across organizational, regional and
national boundaries, to provide timely and seamless portability of information
and optimize the health of individuals and populations globally. Health data
exchange architectures, application interfaces and standards enable data to be
accessed and shared appropriately and securely across the complete spectrum of
care, within all applicable settings and with relevant stakeholders, including
by the individual \citep{HIMSS2019}.

@@

@@

\begin{defn}
A quality of the software production process, referring to its efficiency and
performance \citep{ghezzi1991fundamentals}.
\end{defn}
\begin{defn}
The best definition of the productivity of a process is
\[\text{Productivity} = \dfrac{\text{Outputs produced by the
      process}}{\text{Inputs consumed by the process}}\] Defining inputs. For
the software process, providing a meaningful definition of inputs is a
nontrivial but generally workable problem. Inputs to the software process
generally comprise labor, computers, supplies, and other support facilities and
equipment. Defining outputs. The big problem in defining software productivity
is defining outputs. Here we find a defining delivered source instructions (DSI)
or lines of code as the output of the software process is totally inadequate,
and they argue that there are a number of deficiencies in using DSI. However,
most organizations doing practical productivity measurement still use DSI as
their primary metric \citep{Boehm1987}.
\end{defn}
\begin{defn}
\label{ProductivitySelected}
Productivity is the amount of output (what is produced) per unit of input
used.If we can measure the size of the software product and the effort required
to develop the product, we have:
\begin{align}
\text{productivity} = \text{size}/\text{effort}
\end{align}
Equation (1) assumes that size is the output of the software production process
and effort is the input to the process. This can be contrasted with the
viewpoint of software cost models where we use size as an independent variable
(i.e., an input) to predict effort which is treated as an output. Equation (1)
is simple to operationalize if we have a single dominant size measure, for
example, product size measured in lines of code \citep{Kitchenham2004}.
\end{defn}

@@

@@

@@

@@

\item Verification - Are we building the product right?  Are we implementing the
  requirements correctly (internal)
\item Validation - Are we building the right product? Are we getting the right
  requirements (external)
\item According to
  \href{https://en.wikipedia.org/wiki/Software_verification_and_validation}{Capability
    Maturity Model (CMM)}
\begin{itemize}
\item 
    Software Verification: The process of evaluating software to determine
    whether the products of a given development phase satisfy the conditions
    imposed at the start of that phase. [IEEE-STD-610]
  \item Software Validation: The process of evaluating software during or at the
    end of the development process to determine whether it satisfies specified
    requirements. [IEEE-STD-610] 
\end{itemize}
\end{itemize}

>>>>

\begin{defn}
  ``The capability of a software product to be installed in a specified
  environment'' \citep{berander2005software}.  \wss{remove the quotation marks -
    we are assuming all definitions are directly quoted from the source.  Please
  remove quotes for all of the definitions that you ``own.''}

@@

\begin{defn}
  The degree to which a system or component facilitates the establishment of
  test criteria and the performance of tests to determine whether those criteria
  have been met \citep{IEEEStdGlossarySET1990}. \wss{The author needs to be
    fixed so that this displays properly.}
\end{defn}

@@

\begin{defn}
  Validatability of a software is the degree of ease in validating (checking)
  that software meets user needs. \wss{Need a citation}
\end{defn}

@@

\subsection{\notdone{Robustness} \pmi{owner}}
\begin{defn}
  The degree to which a system or component can function correctly in the
  presence of invalid inputs or stressful environmental conditions
  \citep{IEEEStdGlossarySET1990}.
\end{defn}
\begin{defn}
  The quality can be further informally refined as the ability of a software to
  keep an acceptable behaviour, expressed in terms of robustness requirements, in
  spite of exceptional or unforeseen execution conditions (such as the
  unavailability of system resources, communication failures, invalid or
  stressful inputs, etc.) \citep{fernandez2005model}.
\end{defn}
\begin{defn} \label{RobustnessPartialDefn}
  Code possesses the characteristic of robustness to the extent that it can
  continue to perform despite some violation of the assumptions in its
  specification \citep{boehm2007software}.
\end{defn}
\begin{defn} \label{RobustnessDefnSelected}
  A program is robust if it behaves ``reasonably'', even in circumstances that
  were not anticipated in the requirements specification - for example, when it
  encounters incorrect input data or some hardware malfunction
  \citep{ghezzi1991fundamentals}.
\end{defn}

@@

Definition \ref{RobustnessDefnSelected} rephrased \wss{Aren't you really
  combining Definition~\ref{RobustnessDefnSelected} with
  Definition~\ref{RobustnessPartialDefn}?} : Software possesses the
characteristic of robustness if it behaves ``reasonably'' in two situations: i)
when it encounters circumstances not anticipated in the requirements
specification; and ii) when the assumptions in its requirements specification
are violated.

@@

Combined definition \ref{PerformanceDefnSelected} and
\ref{PerformanceDefnSelected2}: The degree to which a system or component
accomplishes its designated functions within given constraints, such as speed
(database response times, for instance), throughput ( transactions per second),
capacity (con-current usage loads), and timing (hard real-time
demands). \wss{What is the difference between speed and timing?}

@@

This definition includes all important categories of performance.  \wss{What are
  the important categories?}

@@

\begin{defn} \label{MaintainabilityDefnSelected1}
The ease with which a software system or component can be modified to correct
faults, improve performance or other attributes, or adapt to a changed
environment \citep{IEEEStdGlossarySET1990}. 
\end{defn}
\begin{defn}
ISO/IEC 25010 refers to maintainability as the degree of effectiveness and
efficiency with which a product or system can be modified by the intended
maintainers \citep{ISO/IEC25010}. \wss{The author needs to be updated so that
  this display properly.}
\end{defn}
\begin{defn}
Effort required to locate and fix an error in a program
\citep{pressman2005software}.
\end{defn}
\begin{defn}
A set of attributes that bear on the effort needed to make specified
modifications (which may include corrections, improvements, or adaptations of
software to environmental changes and changes in the requirements and functional
specifications) \citep{pfleeger2006software}.
\end{defn}
\begin{defn}
We will view maintainability as two separate qualities: repairability and
evolvability. Software is repairable if it allows the fixing of defects; it is
evolvable if it allows changes that enable it to satisfy new requirements
\citep{ghezzi1991fundamentals}.
\end{defn}
\begin{defn} \label{MaintainabilityDefnSelected2}
Code possesses the characteristic of maintainability to the extent that it
facilitates updating to satisfy new requirements or to correct deficiencies
\citep{boehm2007software}.
\end{defn}

\begin{defn}
  The effort required to locate and fix an error in an operational program
  \citep{McCallEtAl1977}. (As summarized in \citet{VanVliet2000}.) \wss{It looks
    like \citet{pressman2005software} was using \citet{McCallEtAl1977} for their
    definition.} \wss{PM - can you please look into whether
    \citet{pressman2005software} cite \citet{McCallEtAl1977}?  Do they give a
    reason for dropping the word operational?}
\end{defn}

@@

This definition includes all potential reasons to modify the software.
\wss{What are the potential reasons?  How do you know that we haven't missed
  any?}  \wss{What are the other attributes?  What is an attribute?  Is it the
  same as a quality?  We are calling performance a quality in this document.}

@@

\begin{defn}
  The extent to which a program (or parts thereof) can be reused in other
  applications \citep{McCallEtAl1977}. (As summarized in
  \citet{VanVliet2000}.)
\end{defn}

@@

\pmi{I think this is done.} \wss{Looks good at this time.}

\subsection{\notdone{Portability} \pmi{owner}}
\begin{defn}
The ease with which a system or component can be transferred from one hardware
or software environment to another \citep{IEEEStdGlossarySET1990}.
\end{defn}
\begin{defn}
An application is portable across a class of environments to the degree that the
effort required to transport and adapt it to a new environment in the class is
less than the effort of redevelopment \citep{mooney1990strategies}.
\end{defn}
\begin{defn} \label{PortabilityDefnSelected}
Effort required to transfer the program from one hardware and/or software system
environment to another \citep{pressman2005software}.
\end{defn}
\begin{defn}
A set of attributes that bear on the ability of software to be transferred from
one environment to another (including the organizational, hardware, of software
environment) \citep{pfleeger2006software}.
\end{defn}
\begin{defn}
Code possesses the characteristic of portability to the extent that it can be
operated easily and well on computer configurations other than its current
one. This implies that special function features, not easily available at other
facilities, are not used, that standard library functions and subroutines are
selected for universal applicability, and so on \citep{boehm2007software}.
\end{defn}
\begin{defn}
Portability refers to the ability to run a system on different hardware
platforms \citep{ghezzi1991fundamentals}.
\end{defn}
\begin{defn}
  ``The effort required to transfer a program from one hardware and/or software
  environment to another'' \citep{McCallEtAl1977}. (As summarized in
  \citet{VanVliet2000}.)
\end{defn}

@@

Definition \ref{PortabilityDefnSelected} rephrased: Effort required to transfer
a program between system environments. \wss{Should we clarify what a system
  environment is?  What is a system environment?  Hardware and OS?}

@@

\begin{defn}
Interoperability is the ability of two or more systems or components to exchange
information and to use the information that has been exchanged
\citep{IEEEComputerDictionary1991}.
\end{defn}
\begin{defn}
\label{InteroperabilitySelected}
The degree to which two or more systems, products or components can exchange
information and use the information that has been exchanged
\citep{ISO/IEC25010}.
\end{defn}
\begin{defn}
The ability of a system to coexist and cooperate with other systems.
\citep{ghezzi1991fundamentals}.
\end{defn}
\begin{defn}
The capability to communicate, execute programs, and transfer data among various
functional units in a manner that requires the user to have little or no
knowledge of the unique characteristics of those units
\citep{ISO/IEC/IEEE24765}. \wss{We need to fix the author so that this displays properly}
\end{defn}
\begin{defn}Interoperability is a characteristic of a product or system, whose
interfaces are completely understood, to work with other products or systems,
present or future, in either implementation or access, without any restrictions
\citep{AFUL2019}.
\end{defn}
\begin{defn}
Interoperability is the ability of different information systems, devices and
applications (‘systems’) to access, exchange, integrate and cooperatively use
data in a coordinated manner, within and across organizational, regional and
national boundaries, to provide timely and seamless portability of information
and optimize the health of individuals and populations globally. Health data
exchange architectures, application interfaces and standards enable data to be
accessed and shared appropriately and securely across the complete spectrum of
care, within all applicable settings and with relevant stakeholders, including
by the individual \citep{HIMSS2019}.

@@

\wss{Did you also look for transparency?  You should be able to find definitions
  for transparency, especially if you look outside of a software context.  For
  instance, \url{https://www.macmillandictionary.com/dictionary/british/transparent}}

@@

\begin{defn}
A quality of the software production process, referring to its efficiency and
performance \citep{ghezzi1991fundamentals}.
\end{defn}
\begin{defn}
The best definition of the productivity of a process is
\[\text{Productivity} = \dfrac{\text{Outputs produced by the
      process}}{\text{Inputs consumed by the process}}\] Defining inputs. For
the software process, providing a meaningful definition of inputs is a
nontrivial but generally workable problem. Inputs to the software process
generally comprise labor, computers, supplies, and other support facilities and
equipment. Defining outputs. The big problem in defining software productivity
is defining outputs. Here we find a defining delivered source instructions (DSI)
or lines of code as the output of the software process is totally inadequate,
and they argue that there are a number of deficiencies in using DSI. However,
most organizations doing practical productivity measurement still use DSI as
their primary metric \citep{Boehm1987}. \wss{Is this a direct quote from
  \citet{Boehm1987}?  The sentences seem incomplete?}
\end{defn}
\begin{defn}
\label{ProductivitySelected}
Productivity is the amount of output (what is produced) per unit of input
used. If we can measure the size of the software product and the effort required
to develop the product, we have:
\begin{align}
\text{productivity} = \text{size}/\text{effort}
\end{align}
Equation (1) assumes that size is the output of the software production process
and effort is the input to the process. This can be contrasted with the
viewpoint of software cost models where we use size as an independent variable
(i.e., an input) to predict effort which is treated as an output. Equation (1)
is simple to operationalize if we have a single dominant size measure, for
example, product size measured in lines of code \citep{Kitchenham2004}.
\end{defn}

@@

It is concise and measurable. \wss{What is the output?  What is the input?  I
  think the definition needs to give more information on these.  In particular,
  the above definitions focus on code as the output, but documentation, test
  cases etc should also be part of the output.  If we are going to measure this,
  we need a better idea of what we are measuring for outputs and inputs.}

@@

\wss{Were there any other definitions of completeness?  You could add the
  definition of completeness from \citet[p.\ 5--6]{IEEE1998}.  This definition is for
  requirements, but maybe there is something we can generalize from the
  definition?  We could also look for definitions outside of software development.}

@@

\wss{The definition from \citet{IEEE1998} might again be useful.  We could also
  look for definitions outside of software development.  Even a dictionary
  definition could be helpful.}

@@

\item Verification - Are we building the product right?  Are we implementing the
  requirements correctly (internal)
\item Validation - Are we building the right product? Are we getting the right
  requirements (external)
\item According to
  \href{https://en.wikipedia.org/wiki/Software_verification_and_validation}{Capability
    Maturity Model (CMM)}
\begin{itemize}
\item 
    Software Verification: The process of evaluating software to determine
    whether the products of a given development phase satisfy the conditions
    imposed at the start of that phase. [IEEE-STD-610] \wss{Need a proper citation}
  \item Software Validation: The process of evaluating software during or at the
    end of the development process to determine whether it satisfies specified
    requirements. [IEEE-STD-610] \wss{Need a proper citation}
\end{itemize}
\end{itemize}
